Assessment of zonally symmetric and asymmetric components of the Southern Annular Mode using a novel approach

### 2021-05-26 analysis/paper/asymsam.Rmd
Cita a thompson
Closes #15

Even though these indices are based on zonal averages, their associated geopotential height spatial anomalies contain noticeable deviations from zonal symmetry, particularly in the Pacific Ocean region.
The zonal asymmetries have not been widely studied, but previous work suggest that they strongly modulate the regional impacts of the SAM [@fan2007; @silvestri2009; @fogt2012; @rosso2018].
The fact that the SAM is not entirely zonally symmetric hinders our ability to reconstruct its historical variability prior to the availability of dense observations in the Southern Hemisphere [@jones2009].

@@

Figure \@ref(fig:regr-air-season) shows regression coefficients of each index at 700 hPa with surface land air and sea temperature for each trimester.
In Summer, positive values of the SAM index (Figure \@ref(fig:regr-air-season)`r get_tag(season == "DJF" & term == "full")`) are associated with negative temperature anomalies near Antarctica which are surrounded by a ring of positive anomalies in the mid latitudes.
The ring is not zonally symmetric, as there are four distinctive local maximums around 30\degree W, 120\degree W, 150\degree E and 90\degree E respectively.
In the tropics, there are negative anomalies in the equatorial Pacific, consistent with the negative correlation between SAM and ENSO.
Figure \@ref(fig:regr-air-season)`r combine_words(get_tag(season == "DJF" & term != "full", n = 2))` show temperature anomalies associated with positive values of A\nobreakdash-SAM and S\nobreakdash-SAM, respectively.
Both the local maximums in the ring and the anomalies in the Pacific regions are present mostly on the A\nobreakdash-SAM regression map, while temperature patterns linked to positive S\nobreakdash-SAM show a more zonally consistent ring and less relation to the tropics.
Over Antarctica positive values of the SAM index are associated with negative temperature anomalies, particularly over the Eastern coast.
These anomalies are associated only with the S\nobreakdash-SAM.
On the other hand, temperature anomalies in the Indian ocean, South Africa and Australia are strongly related to A\nobreakdash-SAM and are not present in the regression pattern with the SAM.

>>>>

Even though these indices are based on zonal averages, their associated geopotential height spatial anomalies contain noticeable deviations from zonal symmetry, particularly in the Pacific Ocean region.
The zonal asymmetries have not been widely studied, but previous work suggests that they strongly modulate the regional impacts of the SAM [@fan2007; @silvestri2009; @fogt2012; @rosso2018].
The fact that the SAM is not entirely zonally symmetric hinders our ability to reconstruct its historical variability prior to the availability of dense observations in the Southern Hemisphere [@jones2009].

@@

Figure \@ref(fig:regr-air-season) shows regression coefficients of each index at 700 hPa with surface land air and sea temperature for each trimester.
In Summer, positive values of the SAM index (Figure \@ref(fig:regr-air-season)`r get_tag(season == "DJF" & term == "full")`) are associated with negative temperature anomalies near Antarctica which are surrounded by a ring of positive anomalies in the mid latitudes.
The ring is not zonally symmetric, as there are four distinctive local maximums around 30\degree W, 120\degree W, 150\degree E and 90\degree E respectively.
In the tropics, there are negative anomalies in the equatorial Pacific, consistent with the negative correlation between SAM and ENSO.
Figure \@ref(fig:regr-air-season)`r combine_words(get_tag(season == "DJF" & term != "full", n = 2))` show temperature anomalies associated with positive values of A\nobreakdash-SAM and S\nobreakdash-SAM, respectively.
Both the local maximums in the ring and the anomalies in the Pacific regions are present mostly on the A\nobreakdash-SAM regression map, while temperature patterns linked to positive S\nobreakdash-SAM show a more zonally consistent ring and less relation to the tropics.
Over Antarctica positive values of the SAM index are associated with negative temperature anomalies, particularly over the Eastern coast.
These anomalies are associated only with the S\nobreakdash-SAM.
Furthermore, the typical longitudinal variations of temperature anomalies along the Antarctic Peninsula are not evident in the regressions with the SAM, in agreement with previous works [e.g. marshall2016].
On the other hand, temperature anomalies in the Indian ocean, South Africa and Australia are strongly related to A\nobreakdash-SAM and are not present in the regression pattern with the SAM.
